---
title: "Analyzing the impacts of temperature on rockfish recruitment in the Channel Islands"
date: 12/4/2025

---

![Photo from SeaDoc Society](images/rf_yoy.pdf){fig-align="center"}

## Background

Rockfish (genus *Sebastes*) are an important commercial and recreational fishing target on the west coast. The early life stages of Rockfish are spent in the water column until they grow strong enough to swim to structure. With global ocean temperatures increasing, I aim to analyze how increases in water temperature during Pelagic Larval Duration (PLD) affects the abundance of new rockfish, or Young of Year (YOY). I will utilize timeseries data of YOY abundance collected at the Northern Channel Islands. During data collection, California experienced a marine heatwave often referred to as "The Blob" during 2014-2016. I will analyze the effects of temperature on YOY abundance from 2012 to 2019 in order to determine if the warmer waters during the Blob influenced abundance and give insight to how the population may change has global ocean temperatures increase due to climate change.

## About the data

The data quantifying Rockfish recruitment comes from the Partnership of Interdiscpilnary Sciences of Coastal Oceans (PISCO) long-term fish recruitment monitoring in the Channel Islands. The YOYs were collected using Standardized Monitoring Units for the Recruitment of Fishes (SMURFs) that were placed just off the coast of the Northern Channel Islands in an attempt to collect recruit fish as they swam to structure. Monitoring took place from 2000-2018, with collections occuring year round starting in 2012.

The data quantifying temperature is Sea Surface Temperature (SST) from NOAA.

## Hypotheses

H0: Temperature has no effect on rockfish recruitment

Ha: Increased temperature has a positive effect on rockfish recruitment

## DAG

![](images/dag_rfyoy.pdf){fig-align="center"}

## Load necessary packages and data

```{r}
library(tidyverse)
library(knitr)
library(pscl)
library(modelsummary)
library(MASS)


smurf <- read_csv("data/PISCO_UCSB_subtidal_recruitment_fish_data.1.2.csv")

smurf_sp <- read_csv("data/PISCO_UCSB_subtidal_recruitment_fish_spp_list.1.2.csv")



# Data for ocean temperature and upwelling
ocean <- read_csv("data/ALL_OCEAN_DATA.csv")


```

## Data Cleaning

Before I can start working on my statistical modelings, I need to do some cleaning and filtering to get the data ready to be analyzed. I first subset my data to include only my species of interest. I chose Copper, Kelp, Gopher, Grass, and Brown rockfish, because they have similar PLDs and are difficult to differentiate at a small size. I chose to look at data between 2012-2019 to include before, during, and after the Blob. I also narrowed the window of sampling by year to only include sampling during the months of recruitment to avoid a seasonal pattern from emerging. I chose to only look at the two most consistently sampled sites, which were locatied at Santa Cruz island.

```{r}
recruits <- left_join(smurf, smurf_sp)


recruits$Date <- as.Date(paste(recruits$year, recruits$month, recruits$day, sep = "/"))
ocean$Date <- as.Date(ocean$Date, format = "%m/%d/%Y")

recruit_ocean <- left_join(recruits, ocean)
  

rf_yoy <- recruit_ocean %>% 
  filter(year >= 2012  &
        (month >= 2 & month <= 7)) %>% 
  filter(str_detect(site_code, "HAZ")| # Filter data for on Pelican and Hazards sample sites
            str_detect(site_code, "PEL")) %>% 
  filter(classcode == "KGBC" | classcode == "SCAR/SCAU" | classcode == "SATR") #KGB - SCAR/SCAU - SATR 
      
  
summarized_yoy <- rf_yoy %>% 
  group_by(classcode, year) %>% 
  summarise(count = sum(total_fish_collected))


highest_rf <- recruits %>% 
  filter(genus == "Sebastes") %>% 
  group_by(classcode) %>% 
  summarise(count = sum(total_fish_collected)) %>% 
  filter(count >= 1000)

most_yoy <- highest_rf$classcode
```

```{r}
rf_yoy_av <- rf_yoy %>% 
  group_by(classcode, year) %>% 
  summarise(av_fish = mean(total_fish_collected)) %>% 
  ungroup()
```

```{r}
ggplot(rf_yoy_av, aes(year, av_fish)) +
  geom_line(aes(color = classcode)) +
  scale_color_manual(values = c("darkgreen", "brown4", "lightblue")) +
  labs(x = "Year",
       y = "Annual total fish collected") +
  theme_light()
```